Comparing Power Optimizations in the Reorder Buffer

### 2011-06-08 paper/paper.tex
Changes

\subsection{Distributed ROB}

@@

\subsubsection{Distributed (Cluster based) ROB}

>>>>

\subsection{Distributed ROB}
An accepted optimization for improving the power consumption of an ROB unit 
is to break up the ROB buffer into clusters REF[XXX].  In this organization, the in-flight 
instruction information is distributed across all of the units and then re-assembled 
back into program order after the fact.  The primary advantage of this approach 
is that it dramatically lessens the size of the individual ROB units.  This reduces overall
power by reducing the bit line width as well as unit capacitance.

...IMAGE FROM OUR SLIDES...

The overhead associated with this scheme is that the addition of a separate FIFO unit
is required to maintain a global state of instruction order between the separate ROB units.
This central FIFO unit is then referenced upon instruction re-order to commit the instructions
in program order.  The contents of the FIFO unit are fairly simple and incur reasonable hardware
overhead.  Each entry needs only contain the PC of the instruction (or some other unique identifier)
and the corresponding ROB unit that houses the placeholder of the instruction as it is in-flight. 
The total ROB bit storage count remains the same from the non-distributed model to the distributed model
as the overall size of the ROB remains the same.  However, the inclusion of the central FIFO queue 
induces some overhead bits that must be considered when analyzing the overall effectiveness of the 
model.

The distribution or clustering of the ROB unit also inherently induces a minimal performance overhead 
in particular cases.  By breaking the ROB into smaller pieces and distributing the instruction 
data across the unit, the placement of instructions becomes a performance factor.  Typically, a 
distributed ROB model features only a single write port per ROB cluster.  As a result only one operand 
can be forwarded at a time from the ROB.  In the rare case that an instruction requires two operands be 
forwarded from the same ROB cluster a delay of one cycle is imposed while the operands are forwarded 
sequentially.  It is reported, however, that this happens very rarely in practice and induces less than 1\% 
performance degradation REF[XXX].
\subsection{Dynamic ROB}
\subsection{Latch ROB}

@@

\subsubsection{Distributed (Cluster based) ROB}
Our distributed (cluster based) ROB is modeled similarly to the circular ROB.  Instead of having a single ROB buffer, 
numerous ROB buffers are instantiated.  In a real system, there would be one ROB buffer for each functional 
unit cluster.  However, since we do not model functional units in our simulator, we simply divide the ROB 
into multiple smaller ROBs.  Each of these ROBs implement the default circular ROB model.  Each ROB segment 
features a head pointer and a tail pointer and contains an equal portion of the total ROB space.  No further 
enhancements need to be made to the default circular ROB model as the inclusion of the central FIFO allows the 
ROB clusters to operate entirely independently of one another.  Each ROB might as well be the only ROB in the system.

The largest modeling change required by the distributed ROB is the inclusion of the central FIFO queue.  This 
FIFO queue is responsible for maintaining a global state of order between instructions in the different ROB buffers. 
The FIFO queue is modeled as a modified circular buffer and each entry contains: 1) PC of instruction, 
2) ROB{\char'137}ID where the instruction placeholder resides, and 3) a valid bit corresponding to the completion state of 
the instruction.  The total number of entries in the FIFO queue is equal to the total entry size of the ROB.  
Each active entry in the FIFO queue correlates with an active entry in one of the ROB clusters.

The distributed ROB requires only minor changes to the basic circular buffer to realize its effectiveness.  Instructions are 
read from the Instruction Queue as usual and inserted into their corresponding ROB entry while updating the corresponding tail 
pointer.  Since there are now multiple ROBs into which instructions can be inserted, a quick modulus is performed on the 
destination register of the instruction to divide the instructions equally between each ROB.  This presents the possibility 
for a worst-case situation where the entire instruction stream is blocked due to improper balancing between ROB clusters.  
This situation would occur very rarely in practice and would induce no form of deadlock or livelock in the system -- 
it would merely affect performance.  Possible avoidance techniques can be researched as future work.

At the same time as inserting the instruction from the IQ into the ROB cluster, an entry must be recorded in the central FIFO 
queue to maintain global program order.  At this point, both the instruction PC and the recently determined ROB{\char'137}ID are 
noted in the central FIFO queue.  The tail pointer of the central FIFO queue is increased accordingly to match.

Upon completion of an instruction, an instruction must be flagged as valid in both the ROB cluster in which it resides as well 
as its corresponding central FIFO queue entry.  The central FIFO queue then directs the appropriate ROB to write the instruction 
output to its write port when all prior instructions have been written so that program order is maintained.  This is the second 
major behavioral difference between the circular ROB and the distributed ROB, as writes and write-order is driven and determined by 
the central FIFO queue instead of the ROB cluster itself.
